### 2018-07-20 chapters/01-introduction/intro.tex
Update introduction with more important information

A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the paper and one or more \gls*{cas} for verification, analysis and visualization, among other tasks. Especially in \gls*{stem} literature, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}. However, since \LaTeX{} (an extension of \TeX) is the standard to write articles in \gls*{stem} and we only focus on mathematical writings in this paper, we categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the past 30 years~\parencites{LATEX:Standard}[559]{DigitalTypo}{Knuth}. \LaTeX{} enables printing of mathematical formulae in a structure similar to handwritten style. For example, consider the Jacobi polynomial~\parencite[18.3 in table 1]{NIST:DLMF}
\begin{equation}\label{eq:P}
P_n^{(\alpha , \beta)}(\cos(a\Theta)),
\end{equation}
where $\alpha, \beta > -1$, and $n$ a nonnegative integer. 
This formula is written in \LaTeX{} as
\begin{equation}\label{eq:P-tex}
\verb|P_n^{(\alpha,\beta)}(\cos(a\Theta))|.
\end{equation}

While \LaTeX{} focuses on displaying mathematics, \gls*{cas} focus on computations and user friendly syntax. Therefore, each system uses different representations and syntax. Hence, a writer needs to constantly translate mathematical expressions from one representation to another and back again. Table~\ref{tab:JacobiP-usecase} shows four different representations for the same formula~(\ref{eq:P}).

\begin{table}[ht]
	\centering
	\begin{tabular}{ll}
		\hline
		Systems & Representations \\
		\hline
		\hline
		Generic \LaTeX\ & \verb|P_n^{(\alpha,\beta)}(\cos(a\Theta))| \\ 
		Semantic \LaTeX\ & \verb|\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}| \\
		\Maple & \verb|JacobiP(n,alpha,beta,cos(a*Theta))| \\ 
		\Mathematica & \verb|JacobiP[n,\[Alpha],\[Beta],Cos[a \[CapitalTheta]]]|\\
		\hline
	\end{tabular}
	\caption{Different representations for the same mathematical formula~(\ref{eq:P}).}
	\label{tab:JacobiP-usecase}
\end{table}

Such translations can become complex, if there are non-obvious distinctions between the used systems. For example, a \gls*{cas} could use a normalized form of the function or of their arguments for better performances. In this translation scenario, multivalued functions are particularly difficult~\parencite{AISC:MultivaluedFunctions}. A \gls*{cas} usually defines so called \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}. Thereby, implementations of multivalued functions are discontinuous. In general, positioning branch cuts follow some conventions. They are mostly defined conveniently, such as a straight line. Technically, branch cut are curves and therefore can be defined in more complex shapes, such as spirals. The position of branch cuts can vary from \gls*{cas} to \gls*{cas}~\parencite{Branches:acot}. Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function. While \Maple{} defines the branch cut at ${[-\infty\iunit, -\iunit]}$, ${[\iunit,\infty\iunit]}$ (figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (figure~\ref{fig:acot-cut2}).

@@

An automatic translation process is desirable. Providing translations for mathematical \LaTeX{} expressions is difficult, because the semantic information is absent. Semantics is the meaning of an expression. Consider the example of the Jacobi polynomial in~(\ref{eq:P}). A scientific reader might be able to conclude information like $P$ indicating the Jacobi polynomial and $n$ is being a non-negative integer. However, for a computer it is difficult to understand the expression~(\ref{eq:P-tex}). Since \LaTeX{} is the de facto standard for mathematical expressions, most \gls*{cas} provide import and export functions for \LaTeX{} expressions\footnote{Maple since version Maple 18, \textit{Produce output suitable for LaTeX 2e}, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;\\
Mathematica, \textit{Generating and Importing TeX}, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;\\
Matlab, \textit{\LaTeX{} form of symbolic expression in MATLAB}, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;\\
SageMath, \textit{Sage Math}, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017}. However, because of the explained difficulties, those functions are very limited. In fact, they only work for expressions with unique semantic information, but fail for more complex expressions, such as our Jacobi polynomial example.

The \gls*{nist} has therefore developed a set of semantic \LaTeX{} macros~\parencite{DLMF:Macros}. Each of these macros is tied with a unique well-defined mathematical object and linked to the corresponding definition in the \gls*{dlmf}~\parencite{NIST:DLMF,NIST:DLMF:Paper,NIST:Handbook}. The semantic macros are used to semantically enhance the \gls*{dlmf}. An outgrowth of the \gls*{dlmf} project is the \gls*{drmf}~\parencite{DRMF:14,DRMF:15}, which defines more semantic macros, especially for \gls*{opsf}. In table~\ref{tab:JacobiP-usecase} we have already seen an example of a semantic macro. We call \LaTeX{} expressions with semantic macros \textit{semantic \LaTeX{}} and without \textit{generic \LaTeX}, respectively.

@@

%Furthermore, note that the translator we will present in this thesis is based on not published code from the \gls*{mlp} project and the \gls*{pom}-tagger. We will also use the mentioned set of \Macro s, which are not published yet as well. Therefore, the current version of the translator is not online available yet. However, we are planning to publish the translator soon on the \gls*{drmf} website\footnote{\url{http://drmf.wmflabs.org}}.

%Moreover, I will use 'we' rather than 'I' in the subsequent chapters of this thesis, since I published and discussed my ideas with others including my advisors Dr.~Moritz Schubotz and Professor Abdou Youssef, the hosts during research visits Howard S.~Cohl and J\"urgen Gerhard, and the students Joon Bang and Kevin Chen.\footnote{Special thanks to M.~Schubotz to give me the permission for using this phrase from his doctoral thesis \parencite{MORITZ:Dis}.}

%\section{Goals of the project}
%In this thesis, we will present an automatic tool for translations between semantic \LaTeX{} and the \gls*{cas} \Maple{} and \Mathematica. The translation tool should be lite, in a sense that a user does not need a whole software package to use the program. It is also a part of the \gls*{drmf} project and should provide interactive translations in the \gls*{drmf}. Furthermore, it should provide additional information about the translation process and present solutions if an appropriate translation is not possible.

%Another goal is the translator's extensibility in order to provide the basis for future work.

%\section{Structure}
%In the following chapters, we will discuss our translator and verification techniques. Prior to explaining implementation details, we need to introduce some basic background information in chapter~\ref{ch:background}. We start this chapter with an overview of related work (section~\ref{sec:related-work}). Multivalued functions and branch cuts, which were already mentioned, will be introduced in section~\ref{sec:math-background}. Besides the mathematical background, we will discuss semantics in \LaTeX{} in section~\ref{sec:semantics}. A brief introduction to grammar in languages and how we will use it to understand mathematical expressions will be given in section~\ref{sec:mlp}. The chapter is closed by some definitions for the following chapters.

%The next chapter~\ref{ch:translator} is the main chapter of the thesis. We will explain implementations and discuss problems and solutions of and for the translator. It starts with an outline of the goals and our approaches in section~\ref{sec:gen-appr}. While we will present implementation details for the forward translations in section~\ref{sec:forward-translation}, we will present the developed backward translation in the following section~\ref{sec:backward-translation}.

%In chapter~\ref{ch:evaluation} we will discuss, how we can verify or validate a translated expression. The chapter presents some approaches, such as round trip tests (section~\ref{sec:round-trip}) and numerical tests (section~\ref{sec:numerical-tests}). Extracted formulae from the \gls*{dlmf} and \gls*{drmf} created a comprehensive test suite for these approaches. Section~\ref{sec:test-summary} gives a summary of the results.

%The results and the current status of the translation tool will be discussed in chapter~\ref{ch:results}. In conclusion, chapter~\ref{ch:conc-future-work} explains some approaches to further improve the explained techniques in future work.

>>>>

A typical workflow of a scientist who writes a scientific publication is to use a \gls*{wp} to write the paper and one or more \gls*{cas} for verification, analysis and visualization. Especially in \gls*{stem} literature, \LaTeX\footnote{Note that technically \LaTeX{} is not a \gls*{wp} (\url{https://www.latex-project.org/about/}, seen 07/2017) like \textit{Microsoft Word}. However, since \LaTeX{} (an extension of \TeX) is the standard to write articles in \gls*{stem} and we only focus on mathematical writings in this paper, we categorize \LaTeX{} as a \gls*{wp} as well.} has become the de facto standard for writing scientific publications over the past 30 years~\parencites{LATEX:Standard}[559]{DigitalTypo}{Knuth}. \LaTeX{} enables printing of mathematical formulae in a structure similar to handwritten style. For example, consider the Jacobi polynomial~\parencite[18.3 in table 1]{NIST:DLMF}
\begin{equation}\label{eq:P}
P_n^{(\alpha , \beta)}(\cos(a\Theta)),
\end{equation}
where $\alpha, \beta > -1$, and $n$ a nonnegative integer. 
This formula is written in \LaTeX{} as
\begin{equation*}
\verb|P_n^{(\alpha,\beta)}(\cos(a\Theta))|.
\end{equation*}

While \LaTeX{} focuses on displaying mathematics, a \gls*{cas} concentrates on computations and user friendly syntax. Especially important for a CAS is to embed unambiguous semantic information within the input. Each system uses different representations and syntax in consequence. Hence, a writer needs to constantly translate mathematical expressions from one representation to another and back again. Table~\ref{tab:JacobiP-usecase} shows four different representations for the same formula~(\ref{eq:P}).

\begin{table}[ht]
	\centering
	\begin{tabular}{ll}
		\hline
		Systems & Representations \\
		\hline
		\hline
		Generic \LaTeX\ & \verb|P_n^{(\alpha,\beta)}(\cos(a\Theta))| \\ 
		Semantic \LaTeX\ & \verb|\JacobiP{\alpha}{\beta}{n}@{\cos@{a\Theta}}| \\
		\Maple & \verb|JacobiP(n,alpha,beta,cos(a*Theta))| \\ 
		\Mathematica & \verb|JacobiP[n,\[Alpha],\[Beta],Cos[a \[CapitalTheta]]]|\\
		\hline
	\end{tabular}
	\caption{Different representations for the same mathematical formula~(\ref{eq:P}). Generic \LaTeX{} is the default \LaTeX{} expression. Semantic \LaTeX{} uses special semantic macros to embed semantic information.}
	\label{tab:JacobiP-usecase}
\end{table}

Translations from generic \LaTeX{} to \gls{cas} are difficult to realize since the semantic information are not explicitely given in the input. The \gls{nist} has created a set of semantic \LaTeX{} macros~\cite{DLMF:Macros}. Each macro ties specific character sequences to a well-defined mathematical object and is linked with the corresponding definition in the \gls{dlmf} or \gls{drmf}. These macros embeds necessary semantic information into \LaTeX{} expressions. One example of such a macro is given table~\ref{tab:JacobiP-usecase} for the semantic \LaTeX{} representation for the Jacobi polynomial. The macros provide isolated access to important parts of the mathematical function, such as the arguments.

Even with embedded semantic information, a translation between the systems can be difficult. A typical example of complex problems occurs for multivalued functions~\parencite{AISC:MultivaluedFunctions}. A \gls*{cas} usually defines \textit{branch cuts} to compute principal values of multivalued functions~\parencite{Maple:Cuts}, which makes the implementation of a theoretically continues function to a discontinued presentation of it. In general, positioning branch cuts follows conventions, but can be positioned arbitrarily in many cases. Communicate and explain the decision of defined branch cuts is a critical point for \gls{cas} and can vary between the systems~\parencite{Branches:acot}. Figure~\ref{fig:acot-cut-compare} illustrates two examples of different branch cut positioning for the inverse trigonometric arccotangent function. While \Maple{} defines the branch cut at ${[-\infty\iunit, -\iunit]}$, ${[\iunit,\infty\iunit]}$ (figure~\ref{fig:acot-cut1}), \Mathematica{} defines the branch cut at ${[-\iunit, \iunit]}$ (figure~\ref{fig:acot-cut2}).

@@

This article presents a new approach for automatic \gls{p2c} and vice versa conversions. Translations from presentational to computational (computational to presentational) systems are called forward (backward) translations. A forward translation is denoted with an arrow with the target system language above the arrow. For example,
\begin{equation*}
t \overset{\langMaple}{\mapsto} c,
\end{equation*}
where $t$ is an expression in the \LaTeX{} language and $c$ is an element of the \Maple{} language $\langMaple$. As we will see later in this article, we need to compare mathematical concepts between systems. This is impossible from a mathematical point of view. Consider the irrational mathematical constant $e$, known as Euler's number. The theoretical construct for this symbol cannot be mathematically equivalent to the value \verb|exp(0)| in \Maple, caused by computational and implementational limitations. Instead of using the term \textit{equivalent}, we introduce a \textit{appropriate} and \textit{inappropriate} translations. We call a translation such as
\begin{equation}\label{eq:cos-def}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|cos(z)|
\end{equation}
as \textit{appropriate}, while a translation such as
\begin{equation}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|sin(z)|
\end{equation}
is called \textit{inappropriate}. Note that it is not always as easy as in this example to decide if a translation is appropriate or not. Therefore, this article also presents several validation techniques to automatically verify if a translation is appropriate or inappropriate. In addition to this terminology, we introduce \textit{direct translations}. Later in the paper, we will explain that a translation from one specific mathematical object to its \textit{appropriate} counterpart in the other system is not always possible. We call a translation to the \textit{appropriate} counterpart \textit{direct}. For example, the translation~(\ref{eq:cos-def}) is \textit{direct}, while a translation to the definition of the cosine function
\begin{equation*}
\verb|\cos@{z}| \overset{\langMaple}{\mapsto} \verb|(exp(I*z)+exp(-I*z))/2|
\end{equation*}
is not a \textit{direct} translation.

@@
